Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-29 manuscript.Rmd
Update CRQA discussion

## Cross-Recurrence Quantification Analysis
### Syria as Target
The significant result of DET and NRLINE for social cohesion with all events
supports our first hypothesis that event activity is coupled with online social
cohesion. The frequency of all events are thus associated with
greater shared trajectories with social cohesion than would be expected by
chance. This could be due to a conflation of the different kinds of events. However, when we specifically target positive events, the coupling with social cohesion is strong, running counter to our second hypothesis. In addition to the significant DET and NRLINE metrics seen in the all-events analysis, maxL and TT are significant for positive events, suggesting a more stable relationship. This may be a reflection of citizens' reaction to events that align with their cause. An act of concession by the government or aid from another country may have led to an increase in morale or feelings of hope, in turn leading to a groundswell of support on Twitter in an attempt to boost morale. By contrast, negative event frequency showed weak coupling, counter to our third hypothesis. Although the significant result for maxL suggests stability, the lack of significant results for other metrics indicate that the maxL metric may point toward a particular period where coupling was strong, although negative events were less coupled for the other points in time. Perhaps during times of intense strife, individuals in Syria focused on real-world action, turning to Twitter to report and rally together. Further text analysis should be conducted to decipher the underlying tone of the tweets.

### Syria as Source and Target
By adding Syria as a source in the data filtering, the event dataset increased in size by `r num_source_target - num_target` samples. These additional observations are times where Syria initiated
an event with another country. Events initiated by Syria but directed internally
within the country were already captured by the target-only filtering. Taking outwardly directed events into account dramatically changed the coupling between all events and social cohesion. There is no longer a significant number of shared trajectories or line lengths, indicating that the cohesion on Twitter is not tied to the frequency of overall daily events. Since source events are at the root of this difference, it is plausible that Syrians were focused on the target events as they are immediately impacting them; considering additional source events then overshadowed the relationship between target events and social cohesion online. In contrast, there was a slight increase in coupling strength between positive events and social cohesion that once again can only be due to the externally directed events. These two time series showed greater stability in their coupling (reflected in L). However, the significant ENTR metric suggests that there is more variability in the coupling, so rather than interpreting these two changes as a fundamental change in coupling between source events and positive events, we interpret this as a breaking-up of the trajectories formed by the target events. Although the coupling between social cohesion with negative events were largely unchanged, there was a slight decrease in coupling strength (reflected in maxL) and an increase in complexity of shared trajectories (reflected in rENTR). Similarly to that of positive events, this change may not be reflective of source events themselves but rather a disruption of the relationship with target events.

While interesting, the underlying meaning or drivers of these changes in CRQA results requires further investigation that incorporates the actual content of the tweets themselves. One possible reason (among many) may be that the findings were spurious. The source events may be covering up the relationship between different kinds of target events and Twitter social cohesion. However, we cannot test this hypothesis with the current time series data; future research should explore the semantic content of this tweet corpus to determine the nature of the socially cohered
language.

>>>>

## Cross-Recurrence Quantification Analysis
### Syria as Target
The significant result of DET and NRLINE for social cohesion with all events supports our first hypothesis that event activity is coupled with online social cohesion. The frequency of all events are thus associated with a greater number and greater length of shared trajectories with social cohesion than would be expected by chance. This could be due to a conflation of the different kinds of events. However, when we specifically target positive events, the coupling with social cohesion is strong, running counter to our second hypothesis. In addition to the significant DET,NRLINE, and LAM metrics seen in the all-events analysis, maxL and TT are significant for positive events, suggesting a more stable relationship. This may be a reflection of citizens’ reactions to events that align with their cause. An act of concession by the government or aid from another country may have led to an increase in morale or feelings of hope, in turn leading to a groundswell of support on Twitter in an attempt to boost morale.

Negative event frequency also showed coupling with social cohesion, as indicated by significant DET, NRLINE, and maxL metrics.. The significant result for maxL again suggests greater stability of coupling with negative events than with all events. Perhaps during times of intense strife, individuals in Syria focused on communication and public awareness, turning to Twitter to report and rally together. Further text analysis should be conducted to decipher the underlying tone of the tweets.

### Syria as Source and Target
By adding Syria as a source in the data filtering, the event dataset increased in size by `r num_source_target - num_target` observations. These additional observations were times in which Syria initiated an event with another country. Events initiated by Syria but directed internally within the country were already captured by the target-only filtering. Taking outwardly directed events into account  changed the coupling for all-event filtering methods. There was no longer a significant number of shared trajectories or line lengths for social cohesion and count of all events. The recurrent points were still more likely to fall in diagonal trajectories (as seen in the significant DET metric), but there were fewer trajectories overall. This indicates that the cohesion on Twitter is not as strongly tied to the frequency of overall daily events when considering Syria as the target country, as they are less likely to share the same intensity values across extended periods of time. Since Syria-as-source events must be at the root of this difference, it is plausible that Syrians were focused on the target events as they are immediately impacting them; integrating  additional source events then overshadowed the relationship between target events and social cohesion online.

In contrast, there was a slight change in coupling between positive events and social cohesion that once again can only be due to the externally directed events. These two time series showed greater stability in their coupling (reflected in L) but lower attractor strength (reflected in maxL). This means that—although the longest trajectory in the target-only analysis was disrupted by the addition of source events—the additional observations resulted in an overall greater average length of trajectories. However, the significant ENTR and rENTR metrics suggest that there was more variability in the coupling. Rather than interpreting these two changes as a fundamental change in coupling between source events and positive events, this may indicate that the trajectories formed by the target events were broken by the additional source events. This would also explain the overall decrease in the number of shared trajectories (reflected in NRLINE).

The coupling between social cohesion with negative events changed dramatically. All previously significant metrics (DET, NRLINE, maxL) became non-significant. Similarly to that of positive events, this change may not be reflective of source events themselves but rather a disruption of the relationship with target events. The negative outward-directed source events may not have been of interest to Syrian Twitter users. Those external events may have then overshadowed the true coupled system between negative target events and virtual social cohesion.

While interesting, the underlying meaning or drivers of these changes in CRQA results requires further investigation that incorporates the actual content of the tweets themselves. One possible reason (among many) may be that the processes of social cohesion of a population in the midst of an uprising are more tightly coupled to the state of the country itself, rather than the externally directed actions taken by the ruling class to manage international relations. For example, we observed coupling between negative events and social cohesion when considering only the target-filtered data, but that coupling disappeared when we added the source-filtered data to our dataset. In other words, the source events—that is, those taken by Syria toward parties outside of Syria (since Syrian-directed, Syrian-initiated actions were already included in the target-only dataset)—may have muddied the relationship between different kinds of target events and Twitter social cohesion—that is, the “singing with one voice” of the Syrian populace as they protested against their government. However, we cannot test this hypothesis with the current time series data; future research should explore the semantic content of this tweet corpus to determine the nature of the socially cohered language.
